---
title: "Predicting the percentage of a student based on the number of study hours using Supervised ML with R Programming Language"
author: "Egwolo Akpere"
date: "2022-12-19"
output: html_document
---
#The work reported in this study explores and discuss extensively on a public dataset gotten from The Spark Foundation. The study was preceded by an elaborate and up-to-date analysis using the R programming language. This is a simple linear regression task as it involves just 2 variables. Data can be found at http://bit.ly/w-data

# Business Question: what will be the predicted score if a student studies for 9.25 hrs/ day?

# Solution
# Install and load our package.
```{r}
library(tidyverse)
```

```{r}
library(readr)
```
# Import our dataset

```{r}
data <- read.csv("C:/Users/HP PC/Downloads/Students Data.csv")
```
# Inspect our dataset.
```{r}
View(data)
head(data)
```

# Apply linear regression by creating a model

```{r}
model<- lm(Scores~Hours,data = data)
model
```
# Plot our data. 
#we do this by installing and loading our ggplot2 package
```{r}
library(ggplot2)
```
# Plotting the distribution of scores
```{r}
ggplot(data= data,aes(x=Hours,y=Scores))+geom_point() + ggtitle('Distribution of Scores')+
geom_point(color='red')
```

#From the graph above, we can clearly see that there is a positive linear relation between the number of hours studied and percentage of score.

# Fit a regression line into our plot
```{r}
ggplot(data= data,aes(x=Hours,y=Scores))+geom_point() + ggtitle('Distribution of Scores')+
  geom_point(color='red')+ geom_smooth(method = "lm", se = FALSE)
```

# Predict the scores of our Data Frame using our model
```{r}
predicted_scores <-predict(model, data = data$Scores)
predicted_scores
```
# Add predicted scores to our Dataframe.
```{r}
data$prediction <- predicted_scores
View(data)
head(data)
```
# Check the accuracy of our model
```{r}
cor.test(data$Scores,predicted_scores)
```
#both data have an extimated corrolation of 0.976. now lets visualize this.
```{r}
ggplot(data= data,aes(x=predicted_scores, y=Scores))+geom_point() + ggtitle('Distribution of Scores')+
  geom_point(color='darkblue')
```

# Split the data into training and test set
```{r message=FALSE, warning=FALSE}
library(caret)
set.seed(123)
training.samples <- data$Scores %>%
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data[training.samples, ]
test.data <- data[-training.samples, ]
```
# Build the model
```{r}
model_train <- lm(Scores ~., data = train.data)
```
# Summarize the model
```{r}
summary(model_train)
```
#In our example, it can be seen that p-value of the F-statistic is < 1.983e-14, which is highly significant. This means that, at least, one of the predictor variables is significantly related to the outcome variable.
# Make predictions
```{r}
predictions <- model_train %>% predict(test.data)
```
# Model performance
# (a) Prediction error, RMSE
```{r}
RMSE(predictions, test.data$Scores)
```
#Root mean square error is 3.8695
# (b) R-square
```{r}
R2(predictions, test.data$Scores)
```
#R-square value is 0.9931 which is very good.

#lastly, we answer our stakeholderas question, i.e.predicting the score if a student studies for 9.25 hrs/ day?

```{r}
X <-data.frame(Hours=9.25)
result<-predict(model,X)
print(result)
```
#if a student studies for 9.25hours, he/she is likely to score 92.91 

#Next we evaluate the model by calculating the Mean Absolute Error between our predicted scores and observed scores. we do this by first installing and loading the Metrics package.
```{r}
library(Metrics)
```
# Calculate the Mean Absolute Error
```{r}
mae(data$Scores, predicted_scores)
```
#MAE for this model is 4.973



